---
title: "Readme"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Case 1


Let's say I have a very complex function `f`. In order to have a cleaner and more understandable code, I have created another function `f1` that is called by this main function. These two function are defined in a `.R` file named `functionDefinition.R`:

```{r echo=FALSE, results='asis'}
cat('```r\n')
cat(paste(readLines('functionsDefinition.R'), "\n"))
cat('```')
```


Now I want to save in a `.rds` file the function and another parameter for later used. So I (naively) do:

```{r}
createSetup <- function(funDefFile, y) {
  source(funDefFile)
  z <- paste(y, y) # some complex stuff
  list(param = z,
       fun = f)
}

mySetup <- createSetup('functionsDefinition.R', "toto")
saveRDS(mySetup, 'savedSetup-1.rds')
```


Then later in **a new R-session** I load my "setup" and call my function:

```r
mySetup <- readRDS("savedSetup-1.rds")
mySetup$fun(10)
```

**Question: What will happen?**


<details><summary>Click to see the anwser</summary>

```{sh}
R -q --vanilla -e '
tryCatch({
mySetup <- readRDS("savedSetup-1.rds")
mySetup$fun(10)
}, error = function(err) {
  message(err)
})
'
```

**It raise an error.**

Indeed by doing this way, I saved in the `mySetup` list the definition of `f` but not `f1` so when I load it back `f1` is not defined.

Note: by default the function `source` evaluate the given file in the global environment.

</details>




## Case 2:

If I specify that the function `source` should evaluate the function definition in the local environment, it should not change anything, I still do not save (explicitly) `f1`:

```{r}
createSetup_2 <- function(funDefFile, y) {
  source(funDefFile, environment())
  z <- paste(y, y) # some complex stuff
  list(param = z,
       fun = f)
}
```

```{r}
mySetup <- createSetup_2('functionsDefinition.R', "toto")
saveRDS(mySetup, 'savedSetup-2.rds')
```

Again I load my new setup:

```r
mySetup <- readRDS("savedSetup-2.rds")
mySetup$fun(10)
```

**Question: What will happen?**


<details><summary>Click to see the anwser</summary>

```{sh}
R -q --vanilla -e '
tryCatch({
mySetup <- readRDS("savedSetup-2.rds")
mySetup$fun(10)
}, error = function(err) {
  message(err)
})
'
```

**It works !**

But I don't understand why...

</details>


Thank you very much.
