---
title: "Readme"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      results = "hold")
```



## Case 1


Let's say I have a very complex function `f`. In order to have a cleaner and more understandable code, I have created another function `f1` that is called by this main function. These two function are defined in a `.R` file named `functionDefinition.R`:

```{r echo=FALSE, results='asis'}
cat('```r\n')
cat(paste(readLines('functionsDefinition.R'), "\n"))
cat('```')
```


Now I want to save in a `.rds` file the function and another parameter for later used. So I (naively) do:

```{r}
createSetup <- function(funDefFile, y) {
  source(funDefFile)
  z <- paste(y, y) # some complex stuff
  list(param = z,
       fun = f)
}

mySetup <- createSetup('functionsDefinition.R', "toto")
saveRDS(mySetup, 'savedSetup-1.rds')
```


Then later in **a new R-session** I load my "setup" and call my function:

```r
mySetup <- readRDS("savedSetup-1.rds")
mySetup$fun(10)
```

**Question: What will happen?**


<details><summary>Click to see the anwser</summary>

```{sh}
R -q --vanilla -e '
tryCatch({
mySetup <- readRDS("savedSetup-1.rds")
mySetup$fun(10)
}, error = function(err) {
  message(err)
})
'
```

**It raise an error.**

Indeed by doing this way, I saved in the `mySetup` list the definition of `f` but not `f1` so when I load it back `f1` is not defined.

Note: by default the function `source` evaluate the given file in the global environment.

</details>




## Case 2:

If I specify that the function `source` should evaluate the function definition in the local environment, it should not change anything, I still do not save (explicitly) `f1`:

```{r}
createSetup_2 <- function(funDefFile, y) {
  source(funDefFile, environment())
  z <- paste(y, y) # some complex stuff
  list(param = z,
       fun = f)
}
```

```{r}
mySetup <- createSetup_2('functionsDefinition.R', "toto")
saveRDS(mySetup, 'savedSetup-2.rds')
```

Again I load my new setup:

```r
mySetup <- readRDS("savedSetup-2.rds")
mySetup$fun(10)
```

**Question: What will happen?**


<details><summary>Click to see the anwser</summary>

```{sh}
R -q --vanilla -e '
tryCatch({
mySetup <- readRDS("savedSetup-2.rds")
mySetup$fun(10)
}, error = function(err) {
  message(err)
})
'
```

**It works !**

Indeed, this code is similar to:

```{r results='markup'}
createFun <- function(y){
  function(x){
    x + y
  }
}

foo10 <- createFun(10)
foo42 <- createFun(42)
y <- 1

foo10(0)
foo42(0)
```

In this case, the functions `foo10` and `foo42` look for `y` in different environments. (*in R the values of free variables are searched for in the environment in which the function was defined*)

Moreover, according to this blog post: [How and why to return functions in R](https://www.r-bloggers.com/2015/04/how-and-why-to-return-functions-in-r/) (section: *The nature of closure driven reference leaks*)

> In R when objects are serialized they save their lexical environment (and any parent environments) up until the global environment. The global environment is not saved in these situations. When a function is re-loaded it brings in new copies of its saved lexical environment chain and the top of this chain is altered to have a current environment as its parent. This is made clearer by the following two code examples:

> Example 1: R closure fails to durably bind items in the global environment (due to serialization hack).


```r
f <- function() { print(x) }
x <- 5
f()
## [1] 5
saveRDS(f,file='f1.rds')
rm(list=ls())
f = readRDS('f1.rds')
f()
## Error in print(x) : object 'x' not found
```

> Example 2: R closure seems to bind items in intermediate lexical environments.


```r
g <- function() {
  x <- 5
  function() {
    print(x)
  }
}
f <- g()
saveRDS(f,file='f2.rds')
rm(list=ls())
f = readRDS('f2.rds')
f()
## [1] 5
```



### Warning

Using such enclosure can make the function to store a lot of unnecessary information depending where they had been defined:

(I have modify the function `f` to print all the variable it has access to, except the global environment, see `functionDefinition_2.R`)

```{r}
createSetup_x <- function(funDefFile, y) {
  
  source(funDefFile, environment())
  
  # next variables are not necessary for later and especially not for `f`
  tempVar_1 <- "a"
  tempVar_2 <- "b"
  tempVar_3 <- "c"
  
  z <- paste(y, tempVar_1, tempVar_2, tempVar_3) # some complex stuff
  list(param = z,
       fun = f)
}
```

```{r}
mySetup <- createSetup_x("functionsDefinition_2.R", "toto")
saveRDS(mySetup, 'savedSetup-x.rds')
```


```{sh}
R -q --vanilla -e '
tryCatch({
mySetup <- readRDS("savedSetup-x.rds")
mySetup$fun(10)
}, error = function(err) {
  message(err)
})
'
```


Here we can see that `f` had saved the values of `funDefFile`, `tempVar_1`,`tempVar_2`, `tempVar_3`, `y` and `z`.

Indeed it had been defined in the function `createSetup_x` in which we can find those variable.


We can avoid that by sourcing in an empty new environment:



```{r}
createSetup_y <- function(funDefFile, y) {
  
  env <- new.env(parent = globalenv())
  source(funDefFile, env)
  
  # next variables are not necessary for later and especially not for `f`
  tempVar_1 <- "a"
  tempVar_2 <- "b"
  tempVar_3 <- "c"
  
  z <- paste(y, tempVar_1, tempVar_2, tempVar_3) # some complex stuff
  list(param = z,
       fun = env$f)
}
```

```{r}
mySetup <- createSetup_y("functionsDefinition_2.R", "toto")
saveRDS(mySetup, 'savedSetup-y.rds')
```


```{sh}
R -q --vanilla -e '
tryCatch({
mySetup <- readRDS("savedSetup-y.rds")
mySetup$fun(10)
}, error = function(err) {
  message(err)
})
'
```







</details>




## Other similar cases

<details><summary>Click to expand</summary>

### Case 3


Now let's get rid of the `source` function:

```{r}
createSetup_3 <- function(funDefFile, y) {
  
  f1 <- function(x) { 
    x^2 
  } 
  
  f <- function(x) { 
    f1(x) + 1 
  } 
  
  z <- paste(y, y) # some complex stuff
  list(param = z,
       fun = f)
}
```

```{r}
mySetup <- createSetup_3('functionsDefinition.R', "toto")
saveRDS(mySetup, 'savedSetup-3.rds')
```

Again I load my new setup:

```r
mySetup <- readRDS("savedSetup-3.rds")
mySetup$fun(10)
```

**Question: What will happen?**


<details><summary>Click to see the anwser</summary>

```{sh}
R -q --vanilla -e '
tryCatch({
mySetup <- readRDS("savedSetup-2.rds")
mySetup$fun(10)
}, error = function(err) {
  message(err)
})
'
```

It also work, but this would be expected because it should behave like case 2.

</details>



### Case 4


Now let's source in a new environment:

```{r}
createSetup_4 <- function(funDefFile, y) {
  env <- new.env()
  source(funDefFile, env)
  
  z <- paste(y, y) # some complex stuff
  list(param = z,
       fun = env$f)
}
```

```{r}
mySetup <- createSetup_4('functionsDefinition.R', "toto")
saveRDS(mySetup, 'savedSetup-4.rds')
```

Again I load my new setup:

```r
mySetup <- readRDS("savedSetup-4.rds")
mySetup$fun(10)
```

**Question: What will happen?**


<details><summary>Click to see the anwser</summary>

```{sh}
R -q --vanilla -e '
tryCatch({
mySetup <- readRDS("savedSetup-4.rds")
mySetup$fun(10)
}, error = function(err) {
  message(err)
})
'
```

It also works

</details>


### Other information


Let's have a look on the objects'/setup files' size and hash:

```{r}
setup1 <- createSetup('functionsDefinition.R', "toto")
setup2 <- createSetup_2('functionsDefinition.R', "toto")
setup3 <- createSetup_3('functionsDefinition.R', "toto")
setup4 <- createSetup_4('functionsDefinition.R', "toto")

setup1_2 <- readRDS('savedSetup-1.rds')
setup2_2 <- readRDS('savedSetup-2.rds')
setup3_2 <- readRDS('savedSetup-3.rds')
setup4_2 <- readRDS('savedSetup-4.rds')

object.size(setup1)
object.size(setup1_2)

object.size(setup2)
object.size(setup2_2)

object.size(setup3)
object.size(setup3_2)

object.size(setup4)
object.size(setup4_2)
```


```{r}
digest::digest(setup1)
digest::digest(setup1_2)
identical(setup1, setup1_2)
all.equal(setup1, setup1_2)

digest::digest(setup2)
digest::digest(setup2_2)
identical(setup2, setup2_2)
all.equal(setup2, setup2_2)


digest::digest(setup3)
digest::digest(setup3_2)
identical(setup3, setup3_2)
all.equal(setup3, setup3_2)

digest::digest(setup4)
digest::digest(setup4_2)
identical(setup4, setup4_2)
all.equal(setup4, setup4_2)
```



```{r}
tools::md5sum('savedSetup-1.rds')
tools::md5sum('savedSetup-2.rds')
tools::md5sum('savedSetup-3.rds')
tools::md5sum('savedSetup-4.rds')

file.info(c('savedSetup-1.rds',
            'savedSetup-2.rds',
            'savedSetup-3.rds',
            'savedSetup-2.rds'))$size
```



```{r}
str(setup1$fun)
str(setup2$fun)
str(setup3$fun)
str(setup4$fun)
```




```{r}
all.equal(setup2, setup3)
all.equal(setup2, setup4)
```

```{r}
all.equal(setup1, setup2)
```

</details>


# session info:




```{r sessionInfo, echo=FALSE}
  options(max.print = 10000)
  if (Sys.info()["sysname"] == "Linux") {
    cat("\nCPU: ")
    cat(unique(system("awk -F': ' '/model name/{print $2}' /proc/cpuinfo", intern = T)))
    cat("\nMemory total size: ")
    cat(as.numeric(system("awk '/MemTotal/ {print $2}' /proc/meminfo", intern = T))*10^(-6), "GB")
  }
  cat("\n\n\nSession information:\n")
  print(sessionInfo(), locale = FALSE)
```
